\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{subcaption}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Flight Test of a Collision Avoidance Neural Network with Run-Time Assurance
%{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
%should not be used}
%\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{Darren Cofer, Ramachandra Sattigeri, \\
Isaac Amundson, Junaid Babar, \\
Saqib Hasan}
\IEEEauthorblockA{\textit{Collins Aerospace} \\
first.last@collins.com}
\and
\IEEEauthorblockN{Eric Smith, \\
Karthik Nukala}
\IEEEauthorblockA{\textit{Kestrel Institute} \\
\{eric.smith, nukala\}@kestrel.edu}
\and
\IEEEauthorblockN{Denis Osipychev, Lucca Timmerman, \\
Dragos Margineantu, James Paunicka, \\
Matthew Moser}
\IEEEauthorblockA{\textit{Boeing} \\
\{first.last\}@boeing.com}
}

\maketitle

\begin{abstract}
Aircraft systems have requirements for airworthiness certification that present barriers to the use
of machine learning technologies such as neural networks. Showing that a component or system is
correct and does no harm through behaviors that were unintended by designers or unexpected by
operators is a critical aspect of the certification process. Our team is developing assurance
technologies that can support the use of machine learning in the design of safety-critical aircraft
systems. These capabilities have been integrated on Boeing’s autonomy demonstrator aircraft to show
that they can provide evidence of correct operation and safety guarantees needed by real aircraft.
We have applied run-time assurance along with formal methods modeling and analysis tools to an
airborne collision avoidance system based on a neural network. This system was demonstrated
in-flight and shown to correctly monitor neural network operation and intervene when needed to
prevent violation of the “stay well clear” safety requirement relative to an intruder aircraft.
\end{abstract}

\begin{IEEEkeywords}
machine learning, run-time assurance
\end{IEEEkeywords}

\input{intro}
% motivation
% aerospace applications
% barriers - requirements, verification, coverage and unintended functionality
% brief description of the demo application and clear statement of our accomplishments

\input{challenges}
% Assurance Challenges
% The assurance challenges for the use of neural networks in safety-critical aircraft applications [DARREN]
% show absence of unintended functionality -- "do no harm"

\input{framework}
% Autonomy Framework [MATT/DENIS]
% The autonomy framework and aircraft used to demonstrate the collision avoidance neural network capability.
% Safety requirements
% Reinforcement learning and the LEC produced for collision avoidance.
% Execute LEC repetitively to generate avoidance trajectory.

\input{approach}
% Run-Time Assurance Architecture, monitors, high-assurance components, ASTM F3269-17 standard
% The run-time assurance architecture developed to guarantee the absence of unintended behavior resulting from the neural network
% - RTA, AADL model [DARREN/ISAAC]
% - Trajectory prediction, SWC assessment [RAM/DARREN]

\input{assurance}
% Assurance Technologies
% The formal methods assurance technologies applied within the architecture,
% including analysis of the architecture design and synthesis of decision logic from a formal specification

\input{agree}
% - AGREE analysis of RTA arch in AADL [JUNAID/ISAAC]

\input{logic}
% - Decision logic, tabular spec and evolution, APT synth and proof [ERIC/KARTHIK]

\input{resolute}
% - Resolute Assurance argument for RTA arch [SAQIB/ISAAC]

\input{results}
% Flight Test Results
% Flight testing, results obtained for various test scenarios, and lessons learned from the demonstration [DARREN/JIM]

\input{conclusion}
% next steps
% other applications?

\subsubsection*{Acknowledgment}
This work was funded by DARPA contract  FA8750-18-C-0099. The views, opinions and/or
findings expressed are those of the author and should not be interpreted as representing
the official views or policies of the Department of Defense or the U.S. Government.

\bibliographystyle{abbrv}
\bibliography{biblio}

\end{document}
